Clustering time series based on probability distributions across temporal granularities

### 2021-11-16 main.Rmd

Cluster characterization is an important element of cluster analysis. @Cook2007-qe provides several methods for characterizing clusters. _Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study. A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines. Changing the axes may reveal patterns or relationships between variables for categorical variables. However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable. Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion. All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse. Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space. The t-SNE embedding will compress data points that are close in high-dimensional space. Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space. The cluster characterization approach varies depending on the distance metric used. Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances. For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.

>>>>

Cluster characterization is an important final stage of a cluster analysis. The primary purpose is to compare the homogeneity within a cluster to the heterogeneity of clusters. This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data. @Cook2007-qe provides visual examples using both tours [@asimov] and parallel coordinate plots [@wegman1990hyperdimensional]. Dimension reduction techniques like principal component analysis [@joliffe], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@maarten-tsne] and linear discrimant analysis (LDA) [@lda] are also useful. <!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines. Changing the axes may reveal patterns or relationships between variables for categorical variables. However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable. Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion. All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse. Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space. The t-SNE embedding will compress data points that are close in high-dimensional space. Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space. The cluster characterization approach varies depending on the distance metric used. Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances. For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots. -->
